---
title: "Competition-Sberbank_Russian_Housing_Market_EDA"
author: "Xi Liang"
date: "7/27/2017"
output: html_document
---
```{r}
library(data.table)
library(dplyr)
library(ggplot2)
library(corrplot)
library(zoo)
```

```{r}
train <- fread("data/train.csv", stringsAsFactors = T)
test <- fread("data/test.csv", stringsAsFactors = T)
macro <- fread("data/macro.csv")
```


```{r}
str(train)
```

### Missing Data
```{r}
# missing value count by columns
missing_val <- sapply(train, function(x) sum(is.na(x))) %>% sort(decreasing = T)
missing_val_df <- data.frame(variable = names(missing_val), missing_val_count = as.numeric(missing_val)/nrow(train))

missing_val_df <- missing_val_df %>% filter(missing_val_count > 0)
ggplot(missing_val_df, aes(x = reorder(variable, -missing_val_count), y = missing_val_count)) +
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

### Data Quality
```{r}
train$state[train$state == 33] <- which.max(table(train$state))
train$build_year[train$build_year == 20052009] <- 2007
```


### Correcting Time Stamp
```{r}
train$timestamp <- as.Date(train$timestamp)
```

### Home Price Distribution
```{r}
summary(train$price_doc)

#housing price distribution in millions
hist(train$price_doc/1000000,breaks = 500)
```

### Home Price Time Series
```{r}
ts_df <- train[,c("timestamp", "price_doc")]
ts_df <- ts(ts_df)
ts_df_seasons <- diff(ts_df, lag = 4)
ts.plot(ts_df_seasons)
```

## House Internal Characteristics
### Correlation of home interior features and home prices
```{r}
interior_fea <- c("full_sq", "life_sq", "num_room", "kitch_sq",
                  "floor", "max_floor", "build_year", "state", "price_doc")

train[,..interior_fea] %>% cor(use = "complete.obs") %>% corrplot()
```

### Area of Home and Number of Rooms

We see that the full square area of the house is highly correlated to the price, we should take closer look
```{r}
ggplot(aes(x = full_sq, y = price_doc), data = train) +
  geom_point(color = "blue")
```

```{r}
#identifying the outlier
outlier_fullsq_index <- which(train$full_sq > 2000)
train[outlier_fullsq_index,]
```

```{r}
#removing the outlier
train <- train[-outlier_fullsq_index,]
ggplot(aes(x = full_sq, y = price_doc), data = train) +
  geom_point(color = "blue")
```

### Room Number
```{r}
table(train$num_room)
```

```{r}
hist(train$num_room, breaks = 20)
```

## Sale Type
```{r}
ggplot(aes(x = price_doc), data = train) +
  geom_density(fill = "blue", color = "blue") +
  facet_grid(~ product_type) +
  scale_x_continuous(trans =  "log")
```

```{r}
train %>% 
  group_by(product_type) %>%
  summarise(median_price = median(price_doc))
```

## Build Year
```{r}
summary(train$build_year)
```

```{r}
table(train$build_year)
```

```{r}
train %>% filter(build_year >= 1691 & build_year < 2018) %>%
  ggplot(aes(x = build_year)) +
  geom_histogram()
```

### Build Yeaer and Price
```{r}
train %>%
  filter(build_year > 1691 & build_year < 2018) %>%
  group_by(build_year) %>%
  summarise(mean_price = mean(price_doc)) %>%
  ggplot(aes(build_year, mean_price)) +
  geom_line() +
  geom_smooth()
```

## Timestamp
```{r}
train$timestamp <- as.Date(train$timestamp)

train %>%
  group_by(timestamp) %>%
  summarise(med_price = median(price_doc)) %>%
  ggplot(aes(x = timestamp, y = med_price)) +
  geom_line(color = "black")
```

### Timestamp vol.
```{r}
train %>%
  group_by(timestamp) %>%
  summarise(n = n()) %>%
  ggplot(aes(x = timestamp, y = n)) +
  geom_bar(stat = "identity")
  
```

## Seasonaal Influence
### By Month
```{r}
train %>%
  mutate(month=month(timestamp)) %>%
    group_by(month) %>% 
    summarize(med_price=median(price_doc)) %>%
    ggplot(aes(x=as.integer(month), y=med_price)) +
    geom_line(color='red', stat='identity') + 
    geom_point(color='red', size=2) + 
    scale_x_continuous(breaks=seq(1,12,1)) + 
    labs(x='Month', title='Price by month of year')
```

## By Year and Month
```{r}
train %>%
  mutate(year_month = as.yearmon(timestamp)) %>%
  group_by(year_month) %>%
  summarise(med_price = median(price_doc)) %>%
  ggplot(aes(x = year_month, y = med_price)) +
  geom_line(color = "black")
```

## House State/Material
```{r}
train %>%
  filter(!is.na(state)) %>%
  ggplot(aes(as.factor(state), log10(price_doc))) +
  geom_jitter(color = "grey", alpha = 0.4) +
  geom_violin(fill = "blue", alpha = 0.8)
```

```{r}
train %>%
  filter(!is.na(state)) %>%
  group_by(state) %>%
  summarise(mean(price_doc))
```

### Material
```{r}
table(train$material)
```

### Price and Material
```{r}
train %>%
  filter(!is.na(material)) %>% 
  ggplot(aes(as.factor(material), log(price_doc))) +
  geom_jitter(color = "grey", alpha = 0.3) +
  geom_violin(fill = 'blue', alpha = 0.7)
```

```{r}
train %>%
  filter(!is.na(material)) %>%
  group_by(state = as.factor(material)) %>%
  summarise(med_price = median(price_doc))
)
```

# Floor of Home
```{r}
ggplot(aes(floor, log(price_doc)), data = train) +
  geom_point(color = "blue", alpha = 0.4) +
  geom_smooth(method = "lm", color = 'red')
```

```{r}
train %>%
  select(id, floor, max_floor) %>%
  filter(floor > max_floor)
```

#Demographic
```{r}
demo_vars <- c('area_m', 'raion_popul', 'full_all', 'male_f', 'female_f', 'young_all', 
               'young_female', 'work_all', 'work_male', 'work_female', 'price_doc')
```

```{r}
corrplot(cor(train[, ..demo_vars], use = "complete.obs"))
```

## Districts
```{r}
train$sub_area %>% unique %>% length
```

```{r}
train %>% 
    mutate(area_km=area_m/1000000, density=raion_popul/area_km) %>%
    select(sub_area, density, price_doc) %>%
  group_by(sub_area) %>%
  summarise(density = median(density), med_price = median(price_doc)) %>%
  ggplot(aes(density, med_price)) +
  geom_point() +
  geom_smooth(method = 'lm', color = 'red')
```

### Sales Based on District
```{r}
train %>%
  group_by(sub_area) %>%
  summarise(dist_sales = n()) %>%
  ggplot(aes(x = reorder(sub_area ,dist_sales), y = dist_sales)) +
  geom_bar(stat = 'identity') +
  coord_flip()
```

## School Dist
```{r}
school_chars <- c('children_preschool', 'preschool_quota', 'preschool_education_centers_raion',
                  'children_school', 'school_quota', 'school_education_centers_raion', 
                  'school_education_centers_top_20_raion', 'university_top_20_raion',
                  'additional_education_raion', 'additional_education_km', 'university_km',
                  'price_doc')

corrplot(cor(train[, ..school_chars], use='complete.obs'))
```

### University: top 20 raion
```{r}
table(train$university_top_20_raion)
```

```{r}
train %>% 
  ggplot(aes (as.factor(university_top_20_raion), price_doc)) +
  geom_jitter(color = 'grey', alpha = 0.2) +
  geom_boxplot(fill = 'pink', alpha = 0.7)
```

```{r}
#district with top 20 universities
train %>%
  filter(university_top_20_raion == 3) %>% 
  select(sub_area) %>% unique
```

# Cultural/Recreational Characterisitics
```{r}
cult_chars <- c('sport_objects_raion', 'culture_objects_top_25_raion', 'shopping_centers_raion',                        'park_km', 'fitness_km', 'swim_pool_km', 'ice_rink_km','stadium_km', 'basketball_km',                   'shopping_centers_km', 'big_church_km','church_synagogue_km', 'mosque_km', 'theater_km',                 'museum_km', 'exhibition_km', 'catering_km', 'price_doc')

corrplot(cor(train[, ..cult_chars], use='complete.obs'))
```

## Price and Sport Objects in Dist
```{r}
train %>%
  group_by(sub_area) %>%
  summarise(sport_objects = mean(sport_objects_raion), med_price = median(price_doc)) %>%
  ggplot(aes(sport_objects, med_price)) +
  geom_point() +
  geom_smooth(method ='lm')
```

## Cultural Objects
```{r}
train %>%
  group_by(culture_objects_top_25) %>%
  summarise(med_price = median(price_doc))
```

## Park
```{r}
train %>%
ggplot(aes(park_km, price_doc)) +
  geom_point()
```

# Infrastructure Features
```{r}
inf_features <- c('nuclear_reactor_km', 'thermal_power_plant_km', 'power_transmission_line_km',
                  'incineration_km','water_treatment_km', 'incineration_km', 'railroad_station_walk_km',                   'railroad_station_walk_min', 'railroad_station_avto_km', 'railroad_station_avto_min',                    'public_transport_station_km', 'public_transport_station_min_walk', 'water_km',                          'mkad_km', 'ttk_km', 'sadovoe_km','bulvar_ring_km', 'kremlin_km', 'price_doc')

corrplot(cor(train[, ..inf_features], use='complete.obs'))
```

```{r}
ggplot(aes(x=kremlin_km, y=price_doc), data=train) +
    geom_point(color='grey') + 
    geom_smooth(method='lm', color='red')
```

#Variable Importance
```{r}
library(caret)

completes <- complete.cases(train)

trControl <- trainControl(method = 'none')

rfmod <- train(price_doc ~ . - id - timestamp - incineration_raion,
               method='rf',
               data=train[completes, ],
               trControl=trControl,
               tuneLength=1,
               importance=TRUE)

varImp(rfmod)
```

```{r}
library(Boruta)
```

```{r}
set.seed(123)
boruta <- Boruta(price_doc ~ . - id - timestamp - incineration_raion,
                 data = train[completes,],
                 doTrace = 2)
```

```{r}
plot(boruta, xlab = "", xaxt = "n")
 lz<-lapply(1:ncol(boruta$ImpHistory),function(i)
boruta$ImpHistory[is.finite(boruta$ImpHistory[,i]),i])

names(lz) <- colnames(boruta$ImpHistory)


Labels <- sort(sapply(lz,median))

axis(side = 1,las=2,labels = names(Labels),
at = 1:ncol(boruta$ImpHistory), cex.axis = 0.7)
```

```{r}
final.boruta <- TentativeRoughFix(boruta)
print(final.boruta)
```
```{r}
index <- apply(final.boruta$ImpHistory, 2, median) > 5
which(index == TRUE)
```

# Train vs Test Data
```{r}

```

